Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-07-27 manuscript/bef_stability/manuscript.Rmd
Review until network building

The effect of species richness and network structure on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effects of species richness and network structure on total biomass, productivity
and stability. However, empirical data on the link between network structure and
ecosystem properties are lacking. We used a dataset of standardized fish
sampling in French stream running from 1995 to 2018. We inferred trophic
interactions thanks to recorded body size and diet literature. We then
investigate the direct and indirect effects of stream size and temperature,
species richness and network structure on total biomass and its temporal
stability. We found that species richness and network structure had a positive
relationship with total biomass while they had respectively a negative and
positive effect on temporal stability. We also found that species richness and
network structure are not independent and that environment had a tremendous
effect on total biomass and its temporal stability notably through its effect on
species richness and network structure. Our study highlights that the effects of
species richness and network structure should be assessed conjointly and that it
opens exciting perspectives on how environment, species richness and network
structure can jointly determine total biomass in natural communities.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and organisms. Indeed species richness generally increases primary productivity
and total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Theoretical models also
predict this pattern in plant communities [@tilman_plant_1997;
@loreau_partitioning_2001] and food webs [@schneider_animal_2016]. Resource use
complementarity and selection effects have been proposed to elucidate the
positive effect of species richness on productivity. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
A positive selection effect arises when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
In contrast, positive [@wu_relationship_2015; @li_relationship_2018],
negative or non significant effects of species richness on total biomass have been found
in natural forests [@szwagrzyk_above-ground_2007; @jacob_productivity_2010;
@seidel_relationship_2013]. Some authors suggest that the lack of resource
complementarity or environmental harshness might be responsible for an absence or a
negative effect of species richness [@loreau_partitioning_2001; @he_global_2013;
@wu_relationship_2015]. Overall, under what conditions and through what
mechanisms we should observe positive relationships between biomass and species
richness in natural settings remains an open question.

Many studies showed that species richness also increases temporal stability of
communities,
i.e. variation through time, often quantified by the temporal CV. Such positive
relationships between species richness and temporal stability have been reported
in experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018]. Theoretical works
have clarified which mechanisms generate this positive relationship between
species richness and temporal stability [@loreau_species_2008;
@thibaut_understanding_2013], (1) species synchrony and (2) species variability. Species
synchrony quantify to what extent the fluctuations of abundances or biomasses
through time are correlated among species. A low synchrony in species
fluctuations, i.e.  asynchrony, is synonymous of compensatory dynamic, where the
decrease in abundance of a species is compensated by the increase of an other.
Asynchrony is generally the main mechanism whereby species richness has a
stabilizing effect on temporal stability [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. ^[Colin asked for a word about $cv_{sp}$ and
stability but my opinion is that it is not necessary and that the depth of mecanisms are
currently fairly explained for the introduction.] Such positive relationships
between speciesrichness and temporal are also observed in natural settings, such
as stream fish communities [@franssen_annual_2011], bird and butterfly
communities [@olivier_urbanization_2020].

@@

When assessing the relationship between community structure and ecosystem
properties in natural communities, it is crucial to consider environmental
factors [@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
Environment affects organism performance directly, and thereby, species
abundances, total biomass [@grace_integrative_2016] and temporal stability
[@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @franssen_annual_2011; @friedpetersen_drivers_2020].
For example, greater ecosystem size might increase the number of available
resources and thus total biomass [@post_ecosystem_2000; @doi_resource_2009].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the the effects of environment and community
structure on total biomass and its stability.

@@

Stream fish community constitutes a good model to study the effects of
network structure on total biomass and its temporal stability. Despite that
fish trophic interactions are difficult to measure in natural conditions, a consequent body
of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Especially in stream communities, including fishes,
@brose_predator_2019 found that the prediction of predator prey trophic
interactions based on body size ratio was far better in stream ecosystems than
in other aquatic and terrestrial ecosystems. Furthermore, the biotic activity in stream
ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties. Lastly, most of the diversity-biomass and diversity-stability
relationships, have been assessed in terrestrial ecosystems
while there is a evident need to test such relationships in a variety of
ecosystem types.

We used the ONEMA dataset consisting in standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. 99 sites were sampled
for more than 10 years spanning on the 1995-2018 period. We inferred
1496 interaction networks, based on body size and food diet. The network
structure was summarized by connectance and mean trophic level. Environment
was characterized by stream size, altitude, slope, temperature, flow
and enrichment. Finally total biomass was measured as median total biomass over
the sampling period while its temporal stability was quantified as the inverse
of the CV.

We found a positive relationship between total biomass and species richness and
network structure  while total biomass and species richness had respectively a
negative and positive effect on temporal stability. We also showed a link
between species richness and network structure. We found that the environment had
a strong effect on total biomass and its temporal stability in particular by 
its effect on species richness and network structure. We then discuss the
possible mechanisms at play to explain those results and their implications on
our knowledge of the effect of network structure on ecosystem properties.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in Supplementary material, part 1.1). For streams
shallower than 0.7m across the entire width, the sampling took place over the
whole width of the streams and was done by foot whereas in streams deeper than
0.7m, the sampling was done over the bank by foot or by boat. The sampling
season spanned from late spring to autumn. During sampling, all the fishes
caught were identified and recorded. For each species, either the body size of
all fish was measured, or a representative sample when the number of fish was
high (Figure S1, supplementary information). The fishes were released at the end
of the sampling. Additionally, the physical characteristics of the river was
recorded during sampling as well as the sampling effort, quantified as the
surface sampled.

@@

We kept only the most frequent of the two sampling protocols in each station and kept
stations that had been sampled ten years or more. The median sampling month was
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$, i.e. there was low heterogeneity of the sampling month
within stations. As the concept of community variability referred to variability
around an equilibrium, which is rarely observed or difficult to characterize in
natural communities, only those stations with no temporal trends in total
biomass were selected (detailed in supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results). We
retained 99 stations distributed over seven hydrographic basins (Figure
\@ref(fig:matmet) B).

@@

The body size of the unmeasured individuals was inferred under the assumption
that species body size followed a normal distribution whose mean and
standard deviation were quantified from the recorded individuals of the sampling
(See supplementary methods, Figure S1). The body mass of each individual was then estimated
using an allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$ being the body mass
in grams, $l$ the body size in millimeters and $i$ the individual), where the
coefficients $0.01$ and $3.03$ comes from two meta-analyses
[@lleonart_removing_2000; @froese_cube_2006]. Species biomass
and total biomass at each sampling was then calculated by summing individual
body masses. To account for the sampling effort, species biomass and total
biomass have been related to the surface sampled, i.e. $gm^{-2}$.
In the further analysis, the total biomass of a station has been defined as the
median of the total biomass over time.

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the biomass of
the station $i$). Then, temporal stability was equal to $CV^{-1} = \mu
\sigma^{-1}$.

In order to describe the mechanisms driving temporal stability of biomass, we
decomposed the $CV$ of total biomass ($CV_{com}$) in synchrony ($\phi$) and
weighted average species $CV$ ($\overline{CV_{sp}}$) components according to
@thibaut_understanding_2013:

@@

Trophic interaction network was inferred from body size and ontogenic diet shift.
Body size is a reasonable approximation widely used to infer trophic
interactions [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan. For example, alvins feed on plankton while adults may feed on
algae or fishes. Accordingly, each fish species was divided in nine body size
classes, each corresponding to a trophic species (Figure S2). A high congruence
between inferred interactions and measured trophic interactions has been found
at trophic species scale but not at the species scale [@jennings_weak_2001].
Additionally to the fish trophic species, we added seven resource nodes present
in ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Figure \@ref(fig:matmet)
C & D). The ontogenic diet database was based on fishbase and literature. Each
fish species had two or three life stages which are delimited by body size
(Figure S2).

@@

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
means that species at higher trophic level are more stable than the species at
lower trophic levels. Strong top-down control can generate this pattern in pure
trophic chain [@barbier_pyramids_2018; @shanafelt_stability_2018] but it is not
clear if this theoretical prediction holds when considering omnivorous species
[@wang_biodiversity_2016] and if other mechanisms can generate the same pattern
[@thibaut_understanding_2013]. Another explanation is that the fishes at higher
trophic levels are not more stable themselves, but rather allow for stronger
top-down control and resource complementarity and thus reduce biomass
fluctuations in lower trophic levels [@rooney_structural_2006;
@neutel_reconciling_2007].

>>>>

The effect of species richness and network structure on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effects of species richness and network structure on total biomass, productivity
and stability. However, empirical data on the link between network structure and
ecosystem properties are lacking. We used a dataset of standardized fish
sampling in French stream running from 1995 to 2018. We inferred trophic
interactions thanks to recorded body size and diet literature. We then
investigate the direct and indirect effects of stream size and temperature,
species richness and network structure on total biomass and its temporal
stability. We found that species richness and average trophic level had a positive
relationship with total biomass while they had respectively a negative and
positive effect on temporal stability. We also found that species richness and
average trophic level are not independent and that environment had a tremendous
effect on total biomass and its temporal stability notably through its effect on
species richness and average trophic level. Our study highlights that the effects of
species richness and average trophic level should be assessed conjointly and that it
opens exciting perspectives on how environment, species richness and network
structure can jointly determine total biomass in natural communities.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and taxons. Indeed species richness generally increases primary productivity
and total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Theoretical models also
predict this pattern in plant communities [@tilman_plant_1997;
@loreau_partitioning_2001] and food webs [@schneider_animal_2016]. Resource use
complementarity and selection effects have been proposed to elucidate the
positive effect of species richness on productivity. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
A positive selection effect arises when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].

In contrast, positive [@wu_relationship_2015; @li_relationship_2018],
negative or non significant effects of species richness on total biomass have been found
in natural forests [@szwagrzyk_above-ground_2007; @jacob_productivity_2010;
@seidel_relationship_2013]. Some authors suggest that the lack of resource
complementarity or environmental harshness might be responsible for an absence or a
negative effect of species richness [@loreau_partitioning_2001; @he_global_2013;
@wu_relationship_2015]. Overall, under what conditions and through what
mechanisms we should observe positive relationships between biomass and species
richness in natural settings remains an open question.

Many studies showed that species richness also increases temporal stability of
communities,
i.e. variation through time, often quantified by the temporal CV. Such positive
relationships between species richness and temporal stability have been reported
in experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018]. Theoretical works
have clarified which mechanisms generate this positive relationship between
species richness and temporal stability [@loreau_species_2008;
@thibaut_understanding_2013], (1) species synchrony and (2) species variability. Species
synchrony quantify to what extent the fluctuations of abundances or biomasses
through time are correlated among species. A low synchrony in species
fluctuations, i.e.  asynchrony, is synonymous of compensatory dynamic, where the
decrease in abundance of a species is compensated by the increase of an other.
Asynchrony is generally the main mechanism whereby species richness has a
stabilizing effect on temporal stability [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. Increasing, respectively decreasing, species
variability results in respectively a decrease and an increase of temporal
stability [@thibaut_understanding_2013]. Positive relationships between species
variability and species richness have been observed in experimental grasslands
[@tilman_biodiversity_2006], thereby decreasing stability. However, positive,
neutral and negative relationships have been found in natural settings
[@thibaut_understanding_2013; @olivier_urbanization_2020]. ^[Colin asked for a
word about $cv_{sp}$ and stability but my opinion is that it
is not necessary and that the depth of mecanisms are currently fairly explained
for the introduction.] Overall, positive relationships between species richness and
temporal stability are also observed in natural settings, such as stream fish communities
[@franssen_annual_2011], bird and butterfly communities
[@olivier_urbanization_2020].

@@

When assessing the relationship between community structure and ecosystem
properties in natural communities, it is crucial to consider environmental
factors [@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
Environment affects organism performance directly, and thereby, species
abundances, total biomass [@grace_integrative_2016] and temporal stability
[@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @franssen_annual_2011; @friedpetersen_drivers_2020].
For example, greater ecosystem size might increase the number of available
resources and thus total biomass [@post_ecosystem_2000; @doi_resource_2009].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the effects of environment and community
structure on total biomass and its stability.

@@

Because their trophic interactions can be infered, stream fish community
constitutes a good model to study the effects of network structure on total
biomass and its temporal stability. Despite that fish trophic
interactions are difficult to measure in natural conditions, a consequent body
of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Especially in stream communities, including fishes,
@brose_predator_2019 found that the prediction of predator prey trophic
interactions based on body size ratio was far better in stream ecosystems than
in other aquatic and terrestrial ecosystems. Furthermore, the biotic activity in
stream ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties. Lastly, most of the diversity-biomass and diversity-stability
relationships, have been assessed in terrestrial ecosystems while there is a
evident need to test such relationships in a variety of ecosystem types.

Our study aims to document the relationships between foodweb structure and total
biomass and its temporal stability in natural communities. To do so, we used the
ONEMA dataset consisting in standardized sampling of fish communities in French
streams [@edeline_ecological_2013]. 99 sites were sampled for more than 10 years
spanning on the 1995-2018 period. We inferred 1496 interaction networks, based
on body size and food diet. We question whether species richness, connectance
and average trophic level increase, decrease or are not related to total biomass
and its temporal stability. We further question whether the species richness,
connectance and average trophic level have antagonistic or synergistic effect on
total biomass and its temporal stability. Finally, we question if 
ecosystem size and average temperature increase total biomass and its temporal
stability.

---
#We used the ONEMA dataset consisting in standardized sampling of fish
#communities in French streams [@edeline_ecological_2013]. 99 sites were sampled
#for more than 10 years spanning on the 1995-2018 period. We inferred 1496
#interaction networks, based on body size and food diet. The network structure
#was summarized by connectance and mean trophic level. Environment was
#characterized by stream size, altitude, slope, temperature, flow and enrichment.
#Finally total biomass was measured as median total biomass over the sampling
#period while its temporal stability was quantified as the inverse of the CV.
#
#We found a positive relationship between total biomass and species richness and
#network structure while total biomass and species richness had respectively a
#negative and positive effect on temporal stability. We also showed a link
#between species richness and network structure. We found that the environment had
#a strong effect on total biomass and its temporal stability in particular by 
#its effect on species richness and network structure. We then discuss the
#possible mechanisms at play to explain those results and their implications on
#our knowledge of the effect of network structure on ecosystem properties.
---

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in Supplementary material, part 1.1). For streams
shallower than 0.7m across the entire width, the sampling took place over the
whole width of the streams and was done by foot whereas in streams deeper than
0.7m, the sampling was done over the bank by foot or by boat. The sampling
season spanned from late spring to autumn. During sampling, all the fishes
caught were identified and recorded. For each species, either the body size of
all fish was measured, or a representative sample when the number of fish was
high (Figure S1, supplementary information). The fishes were released at the end
of the sampling. Additionally, the physical characteristics of the river was
recorded during sampling as well as the sampling effort, quantified as the
surface sampled (range XXX-XXX).

@@

The stations where the stream was of intermediate size could be sampled
alternatively with one of the two sampling protocols. When it was the case, we
kept only the most frequently used sampling protocol. Finally, we kept stations
that had been sampled ten years or more. there was low heterogeneity of the sampling month
within stations, with the median sampled month being 
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$. As the concept of community variability referred to variability
around an equilibrium, which is rarely observed or difficult to characterize in
natural communities, only those stations with no temporal trends in total
biomass were selected (detailed in supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results). These
selections resulted in 99 stations distributed over seven hydrographic basins
(Figure \@ref(fig:matmet) A).

@@

The body size of all fish individuals was not recorded when they were too
numerous. The body size of the unmeasured individuals was inferred under the
assumption that species body size followed a normal distribution whose mean and
standard deviation were quantified from the recorded individuals of the sampling
(See supplementary methods, Figure S1). The body mass of each individual was
then estimated using an allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$
being the body mass in grams, $l$ the body size in millimeters and $i$ the
individual), where the coefficients $0.01$ and $3.03$ comes from two
meta-analyses [@lleonart_removing_2000; @froese_cube_2006]. Species biomass and
total biomass at each sampling event was then calculated by summing individual body
masses. To account for the variation in sampling effort among sites, species
biomass and total biomass have been related to the surface sampled, i.e.
$gm^{-2}$. In the further analysis, the total biomass of a station has been
defined as the median of the total biomass over time.

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the annual total
biomass of the station $i$). Then, temporal stability was equal to $CV^{-1} =
\mu \sigma^{-1}$.

In order to describe the mechanisms driving the temporal stability of biomass, we
decomposed the $CV$ of total biomass ($CV_{com}$) in synchrony ($\phi$) and
weighted average species $CV$ ($\overline{CV_{sp}}$) components according to
@thibaut_understanding_2013:

@@

Trophic interaction network was inferred from body size and ontogenic diet shift
following previous studies [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan and thereby over body size. For example, alvins feed on plankton while adults may feed on
algae, fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to a
trophic species (Figure S2), thereby increasing the accuracy of trophic
interactions inference [@jennings_weak_2001]. HERE
Additionally to the fish trophic species, we added seven resource nodes present
in ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Figure \@ref(fig:matmet)
C & D). The ontogenic diet database was based on fishbase and literature. Each
fish species had two or three life stages which are delimited by body size
(Figure S2).

@@

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
means that species at higher trophic level are more stable than the species at
lower trophic levels. Strong top-down control can generate this pattern in pure
trophic chain [@barbier_pyramids_2019; @shanafelt_stability_2018] but it is not
clear if this theoretical prediction holds when considering omnivorous species
[@wang_biodiversity_2016] and if other mechanisms can generate the same pattern
[@thibaut_understanding_2013]. Another explanation is that the fishes at higher
trophic levels are not more stable themselves, but rather allow for stronger
top-down control and resource complementarity and thus reduce biomass
fluctuations in lower trophic levels [@rooney_structural_2006;
@neutel_reconciling_2007].
